Linear Variation and an Optimization of Algorithms for Connected Components Labeling in a Binary Image

### 2016-02-08 paper/paper.tex

The structure is commonly called Union-Find, as the two essential operations
it is supposed to support are Union, which is used to merge some two of the
disjoint sets in one, and Find, which is used to find the representative element
of the set some element belongs to, so that we can always tell if two elements are
in the same set.
The structure can be implemented so that the amortized time complexity for each
operation is $O(\alpha(HW))$, where $\alpha$ is the inverse Ackermann function,
and $\alpha(HW) < 5$ for all remotely practical image sizes. % TODO links

@@

The idea is to run through each row twice: first time to establish label equality
classes within row, and the second time to set unique equal label for all pixels of
same CC.
This gives us the opportunity to maintain just $O(W)$
different labels at each moment, and so the Union-Find structure can be allocated
for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.

The exact maximum number of label equality classes between two adjacent rows
depends on whether the given image is condensed.
For non-condensed images $W=2M$, and maximum number of label equality classes
in two adjacent rows is $M=\frac{W}2$, as two object pixels are both in adjacent
rows and in adjacent columns, then they are neighbors in terms of 8-connectivity.

On the other hand, for condensed images $W=M$, and maximum number of label equality
class in two adjacent rows of cells is $2W=2M$.
See \autoref{fig:suf2:extremal} for illustration on the extremal cases.

@@

This modification of SUF was implemented to compare the optimization we present with
another optimization suggested by Wu, Otoo and Suzuki, % TODO reference
namely decision tree.
Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
applicable for condensed image, as it exploits similar observations on 8-connectivity.

@@

For memory-tight applications such as embedded systems it can be useful to
utilise the suggested condensation in a lazy way, so that without actually
constructing the condensed image we use a function (``view'') calculating color of
each cell with 4 lookups to the original binary image.
This is probably most interesting combined with SUF2 algorithm, as it is a way
to win some time without increasing memory footprint.
However this lazy approach increases the number of image lookups.
So, according to our benchmark, this combination is efficient only for dense
($>30\%$) noise.

@@

\newcommand{\inputtable}[2]{
  \begin{figure}
    \centering
    \includegraphics[height=0.45\textheight, keepaspectratio]{pics/#1x#2.eps}
    \caption{Average CPU time per run on noise images $#1 \times #2$ px}
    \label{fig:noise#1x#2}
  \end{figure}
  % \begin{table}[p]
    % \centering
    % \input{../bench-out/#1x#2}
    % \caption{Average milliseconds of cpu time per run on noise images $#1 \times #2$ px}
    % \label{tab:noise#1x#2}
  % \end{table}
}

% \inputtable{200}{300}
% \inputtable{400}{600}
\inputtable{800}{1200}
% \inputtable{1200}{1800}
\inputtable{2000}{3000}
\inputtable{4000}{6000}
\inputtable{5000}{7500}

@@

\bibitem {sterzh}
Стержанов, М.:
Методики выделения связных компонент в штриховых бинарных изображениях.
Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. 2010, с. 169-174.

>>>>

The structure is commonly called Union-Find, as the two essential operations
it is supposed to support are Union, which is used to merge some two of the
disjoint sets in one, and Find, which is used to find the representative element
of the set some element belongs to, so that we can always tell if two elements are
in the same set.
The structure can be implemented so that the amortized time complexity for each
operation is $O(\alpha(HW))$, where $\alpha$ is the inverse Ackermann function,
and $\alpha(HW) < 5$ for all remotely practical image sizes\cite{CLRS}. % TODO links

@@

The idea is to run through each row twice: first time to establish label equivalence
classes within row, and the second time to set unique equal label for all pixels of
same CC.
This gives us the opportunity to maintain just $O(W)$
different labels at each moment, and so the Union-Find structure can be allocated
for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.

The exact maximum number of label equivalence classes between two adjacent rows
depends on whether the given image is condensed.
For non-condensed images $W=2M$, and maximum number of label equivalence classes
in two adjacent rows is $M=\frac{W}2$, as if two object pixels are both in adjacent
rows and in adjacent columns, then they are neighbors in terms of 8-connectivity.

On the other hand, for condensed images $W=M$, and maximum number of label equivalence
class in two adjacent rows of cells is $2W=2M$.
See \autoref{fig:suf2:extremal} for illustration on the extremal cases.

@@

This modification of SUF was implemented to compare the optimization we present with
another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo},
namely decision tree.
Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
applicable for condensed image, as it exploits similar observations on 8-connectivity.

@@

For memory-tight applications such as embedded systems it can be useful to
utilise the suggested condensation in a lazy way, so that without actually
constructing the condensed image we use a function (``view'') calculating color of
each cell with 4 lookups to the original binary image.
This is probably most interesting when combined with SUF2 algorithm, as it is a way
to win some time while avoiding memory footprint increase.
However this lazy approach increases the number of image lookups.
% So, according to our benchmark, this combination is efficient only for dense
% ($>30\%$) noise.

@@

\newcommand{\inputtable}[2]{
  \begin{figure}
    \centering
    \includegraphics[height=0.45\textheight, keepaspectratio]{pics/#1x#2.eps}
    \caption{Average CPU time per run on noise images $#1 \times #2$ px}
    \label{fig:noise#1x#2}
  \end{figure}
  \begin{table}[p]
    \centering
    \input{tables/#1x#2}
    \caption{Average CPU time per run on noise images $#1 \times #2$ px}
    \label{tab:noise#1x#2}
  \end{table}
}

% \inputtable{200}{300}
\inputtable{400}{600}
\inputtable{800}{1200}
\inputtable{1200}{1800}
\inputtable{2000}{3000}
\inputtable{4000}{6000}
% \inputtable{5000}{7500}

@@

\bibitem {sterzh}
Стержанов, М.:
Методики выделения связных компонент в штриховых бинарных изображениях.
Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. с. 169-174. (2010)

\bibitem {CLRS}
Cormen, H., Leiserson, E., Rivest, L., Stein, C.:
Introduction to Algorithms (Second ed.), MIT Press (2001)

\bibitem {wuotoo}
Wu, K, Otoo, E., Suzuki, K.:
Optimizing two-pass connected-component labeling algorithms.
Pattern Anal. Appl. (2008)
